---
title: "Benchmarks: Logloss"
output:
  html_notebook:
    collapsed: no
    theme: united
    toc: yes
    toc_depth: 1
    toc_float: yes
---

This notebook is all about benchmarking some R code used in this package.

Hardware / Software used:

* Intel i7-4600U
* Compilation flags for C/C++: `-O2 -Wall $(DEBUGFLAG) -mtune=core2` (R's defaults)
* Windows Server 2012 R2
* R 3.4.0 + gcc 7.1

# Libraries

```{r init}
library(data.table)
library(microbenchmark)
library(Rcpp)
library(ggplot2)
```

```{r based}

# Helper function to print data well in tables
print_well <- function(data, digits = 6) {
  
  # To milliseconds
  data <- data / 1000000
  
  # Sprintf helper
  sprintf_helper <- paste0("%.0", digits, "f")
  
  cat("| Min | 25% | 50% | 75% | Max | Mean |  \n| --: | --: | --: | --: | --: | --: |  \n| ", sprintf(sprintf_helper, min(data)), " | ", sprintf(sprintf_helper, quantile(data, probs = 0.25)), " | ", sprintf(sprintf_helper, median(data)), " | ", sprintf(sprintf_helper, quantile(data, probs = 0.75)), " | ", sprintf(sprintf_helper, max(data)), " | ", sprintf(sprintf_helper, mean(data)), " |  \n", sep = "")
  
  return(data)
  
}

# Test case function
# Arguments renamed to avoid recursive clash
test_case <- function(f, preds, labels) {
  cat("Test case: ", do.call(f, list(preds = preds[1:50],
                                     labels = labels[1:50])), "  \n", sep = "")
}

```

# Benchmarking

From a regression vector of 1,000,000 observations:

* Vector A of length=(1000000)
* Vector B of length=(1000000) with 2 classes

```
A = [1, 2, 3, 4, ..., 1000000]
B = [1, 2, 3, 4, ..., 1000000]
```

Get the following Vector C:

```
C = smape(C, B)
```

# Initialize data

```{r bench1}

# How many digits for benchmarking in milliseconds
my_digits <- 6L

# How many runs for benchmarking?
my_runs <- 1000L

# How many observations?
my_obs <- 1000000L

# Generate random data
set.seed(11111)
data <- runif(my_obs, 0, 1)
labels <- round(runif(my_obs, 0, 1), digits = 0)

```

# Benchmarks

```{r bench2, results="asis"}

faster1 <- function(preds, labels) {
  zeroes <- abs(labels) + abs(preds)
  divide <- abs(labels - preds) / zeroes
  divide[zeroes == 0] <- 0
  return(2 * mean(divide))
}
test_case(faster1, preds = data, labels = labels)
data1 <- print_well(microbenchmark(faster1(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 2 =====
faster2 <- function(preds, labels) {
  zeroes <- abs(labels) + abs(preds)
  divide <- abs(labels - preds) / zeroes
  divide[zeroes == 0] <- 0
  return(2 * sum(divide) / length(divide))
}
test_case(faster2, preds = data, labels = labels)
data2 <- print_well(microbenchmark(faster2(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 3 =====
faster3 <- function(preds, labels) {
  zeroes <- abs(labels) + abs(preds)
  divide <- abs(labels - preds) / zeroes
  divide[!zeroes] <- 0
  return(2 * sum(divide) / length(divide))
}
test_case(faster3, preds = data, labels = labels)
data3 <- print_well(microbenchmark(faster3(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 4 =====
faster4 <- function(preds, labels) {
  zeroes <- abs(labels - preds) / ((abs(labels) + abs(preds)) / 2)
  zeroes[is.na(zeroes)] <- 0
  return(sum(zeroes) / length(zeroes))
}
test_case(faster4, preds = data, labels = labels)
data4 <- print_well(microbenchmark(faster4(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 5 =====
faster5 <- function(preds, labels) {
  zeroes <- abs(labels - preds) / (abs(labels) + abs(preds))
  zeroes <- sum(zeroes, na.rm = T)
  return(zeroes * 2 / length(labels))
}
test_case(faster5, preds = data, labels = labels)
data5 <- print_well(microbenchmark(faster5(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 6 =====
faster6 <- function(preds, labels) {
  zeroes <- abs(labels - preds)
  zeroes <- zeroes / (abs(labels) + abs(preds))
  zeroes <- sum(zeroes, na.rm = T)
  return(zeroes * 2 / length(labels))
}
test_case(faster6, preds = data, labels = labels)
data6 <- print_well(microbenchmark(faster6(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 7 =====
cppFunction("double faster7(NumericVector preds, NumericVector labels) {
  NumericVector zeroes = abs(labels) + abs(preds);
  NumericVector divide = abs(labels - preds) / zeroes;
  divide[zeroes == 0] = 0;
  double loss = 2 * sum(divide) / divide.size();
  return(loss);
}")
test_case(faster7, preds = data, labels = labels)
data7 <- print_well(microbenchmark(faster6(data, labels), times = my_runs)$time, digits = my_digits)

# ===== BLOCK 8 =====
cppFunction("double faster8(NumericVector preds, NumericVector labels) {
  int labels_size = labels.size();
  NumericVector zeroes(labels_size);
  zeroes = (abs(labels - preds)) / (abs(labels) + abs(preds));
  LogicalVector nan = is_nan(zeroes);
  double loss = 0;
  for (int i = 0; i < labels_size; i++) {
    if (!nan[i])
      loss += zeroes[i];
  }
  return(2 * loss / labels_size);
}")
test_case(faster8, preds = data, labels = labels)
data8 <- print_well(microbenchmark(faster8(data, labels), times = my_runs)$time, digits = my_digits)

```

# Summary Results

```{r bench3}

data_time <- data.table(rbindlist(list(data.frame(Time = data1, Bench = "faster1"),
                                       data.frame(Time = data2, Bench = "faster2"),
                                       data.frame(Time = data3, Bench = "faster3"),
                                       data.frame(Time = data4, Bench = "faster4"),
                                       data.frame(Time = data5, Bench = "faster5"),
                                       data.frame(Time = data6, Bench = "faster6"),
                                       data.frame(Time = data7, Bench = "faster7"),
                                       data.frame(Time = data8, Bench = "faster8"))))
data_time <- data_time[, t_mean := mean(Time), by = Bench]
data_time <- data_time[, t_median := median(Time), by = Bench]
data_time$Benchs <- data_time$Bench 
levels(data_time$Benchs) <- paste0("faster", 1:8, "= [", sprintf(paste0("%.0", my_digits, "f"), data_time[, list(min(Time)), by = Bench]$V1), ", ", sprintf(paste0("%.0", my_digits, "f"), data_time[, list(max(Time)), by = Bench]$V1), "], mean=", sprintf(paste0("%.0", my_digits, "f"), data_time[, list(mean(Time)), by = Bench]$V1), ", median=", sprintf(paste0("%.0", my_digits, "f"), data_time[, list(median(Time)), by = Bench]$V1))

my_time <- data_time[, list(min(Time), quantile(Time, probs = 0.25), median(Time), quantile(Time, probs = 0.75), max(Time), mean(Time)), by = Bench]
colnames(my_time) <- c("Function", "Min", "25%", "50%", "75%", "Max", "Mean")
my_time <- my_time[order(Mean, decreasing = FALSE), ]
print(my_time, digits = 6)

```

# Plot Results

```{r bench4, fig.height=9, fig.width=10}

ggplot(data = data_time, aes(x = Time)) + geom_histogram(aes(y = ..density..), bins = 20, color = "darkblue", fill = "lightblue") + facet_wrap(~ Benchs, ncol = 2) + geom_vline(aes(xintercept = t_mean), color = "blue", linetype = "dashed", size = 2) + geom_vline(aes(xintercept = t_median), color = "red", linetype = "dashed", size = 2) + labs(x = "Time (Milliseconds)", y = "Density") + theme_bw()

```

```{r bench5, fig.height=6, fig.width=10}
ggplot(data = data_time[, .(Time, Bench)], aes(x = Time, y = ..count.., fill = Bench)) + geom_histogram(aes(y = ..density..), bins = 100, position = "fill") + labs(x = "Time (Milliseconds)", y = "Density") + theme_bw()
```

```{r bench6, fig.height=6, fig.width=10}
ggplot(data = data_time[, .(Time, Bench)], aes(x = Time, y = ..count.., fill = Bench)) + geom_density(position = "fill") + labs(x = "Time (Milliseconds)", y = "Density") + theme_bw()
```

```{r bench7, fig.height=6, fig.width=10}

data_time$MilObs <- (1000 / data_time$Time) * my_obs / 1000000
ggplot(data_time[, .(Bench, MilObs)], aes(x = Bench, y = MilObs, fill = Bench)) + geom_boxplot() + labs(x = "Benchmark", y = "Throughput (Million Obs./s)") + theme_bw()

```

# Scaling Benchmarks

```{r bench_scale1}

Benchmarker <- function(f, size, runs, digits, name) {
  
  data_runs <- list()
  
  for (i in 1:length(size)) {
    set.seed(11111)
    data <- runif(size[i], 0, 1)
    labels <- round(runif(size[i], 0, 1), digits = 0)
    cat("  \n  \n## ", name, " run: ",format(size[i], big.mark = ",", scientific = FALSE), " samples (", format(runs[i], big.mark = ",", scientific = FALSE), " times)  \n  \n", sep = "")
    test_case(f, preds = data, labels = labels)
    cat("  \n")
    data_runs[[i]] <- print_well(microbenchmark(f(data, labels), times = runs[i])$time, digits = digits)
    data_runs[[i]] <- data.table(Bench = as.factor(paste0("[", i, "] ", format(size[i], big.mark = ",", scientific = FALSE))), Function = as.factor(name), Time = data_runs[[i]])
    gc(verbose = FALSE)
  }
  
  return(data_runs)
  
}

bench_size <- c(100, 1000, 10000, 100000, 1000000, 10000000, 100000000)
bench_runs <- c(100000, 50000, 10000, 5000, 1000, 500, 100)

```

```{r bench_scale2, results="asis"}

run1 <- Benchmarker(faster6, bench_size, bench_runs, my_digits, "Pure R")
run2 <- Benchmarker(faster8, bench_size, bench_runs, my_digits, "Rcpp")

```

# Scaling Results

```{r bench_scale3}

run1_all <- rbindlist(run1)
run2_all <- rbindlist(run2)
run_all <- rbind(run1_all, run2_all)
run_all$Repeats <- rep(inverse.rle(list(lengths = bench_runs, values = bench_size)), 2)
run_all$MilObs <- (1000 / run_all$Time) * run_all$Repeats / 1000000
run_time <- run_all[, list(quantile(Time, probs = 0.05), median(Time), quantile(Time, probs = 0.95), mean(Time)), by = list(Function, Bench)]
colnames(run_time) <- c("Function", "Benchmark", "5%", "50%", "95%", "Mean")
run_time$`Mil.Obs/s` <- (1000 / run_time$Mean) * bench_size / 1000000
run_time$`5%` <- format(run_time$`5%`, digits = 6, scientific = FALSE)
run_time$`50%` <- format(run_time$`50%`, digits = 6, scientific = FALSE)
run_time$`95%` <- format(run_time$`95%`, digits = 6, scientific = FALSE)
run_time$Mean <- format(run_time$Mean, digits = 6, scientific = FALSE)
run_time$`Mil.Obs/s` <- format(run_time$`Mil.Obs/s`, digits = 6, scientific = FALSE)

print(run_time[1:(nrow(run_time) / 2)])
print(run_time[(nrow(run_time) / 2 + 1):nrow(run_time)])

```

```{r bench_scale4, fig.height=6, fig.width=10}

ggplot(run_all, aes(x = Bench, y = Time, color = Function, fill = Bench)) + geom_boxplot() + scale_y_log10(labels = scales::comma, breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) + stat_summary(fun.y = mean, geom = "line", aes(group = Function)) + stat_summary(fun.y = mean, geom = "point", aes(group = Function)) + labs(x = "Benchmark", y = "Time (Milliseconds)") + theme_bw()

```

```{r bench_scale5, fig.height=6, fig.width=10}

ggplot(run_all, aes(x = Bench, y = MilObs, color = Function, fill = Bench)) + geom_boxplot() + scale_y_log10(labels = scales::comma, breaks = c(1, 2.5, 5, 7.5, 10, 12.5, 15, 17.5, 20, 22.5, 25), limits = c(1, NA)) + stat_summary(fun.y = mean, geom = "line", aes(group = Function)) + stat_summary(fun.y = mean, geom = "point", aes(group = Function)) + labs(x = "Benchmark", y = "Throughput (Million Obs./s)") + theme_bw()

```

```{r bench_scale6, fig.height=6, fig.width=10}

ggplot(data = run_all, aes(x = MilObs, color = Function, fill = Function, group = Function)) + coord_flip() + stat_ecdf(aes(ymin = ..y.., ymax = 1), alpha = 0.5, geom = "ribbon") + stat_ecdf(geom = "line", size = 2, alpha = 0.75, pad = FALSE) + labs(x = "Throughput (Million Obs./s)", y = "Percentile") + facet_wrap(~ Bench, dir = "h", ncol = 2, scales = "free") + theme_bw()

```

